---
title: "Sparklyr Introduction"
output: html_notebook
---

In this notebook we walk through some basic examples of how to use sparklyr in R. This notebook is based on the guide provided at http://spark.rstudio.com/. 

## Connecting to Spark

Before we start we must install a local instance of Spark
```{r}
library(sparklyr)
spark_install(version = "2.1.0")
```

We can connect to clusters and local instances of Spark as follows (I had to create a new Spark connection first):
```{r}
sc <- spark_connect(master = "local")
```

## Using dplyr

Next we will show that we cn use all of the standard dplyr verbs against tables in the cluster. First we must populate the cluster with data:
```{r}
library(dplyr)
iris_tbl <- copy_to(sc, iris)
flights_tbl <- copy_to(sc, nycflights13::flights, "flights")
batting_tbl <- copy_to(sc, Lahman::Batting, "batting")
src_tbls(sc)
```
Visiting the SparkUI we see that this is a relatively small database (only about 50MB). We could imagine in real cases we would be working with larger databases. Note that the data tables lie `iris_tbl` are lists and are not stored in memory.

Let's try a simple filtering example:
```{r}
# filter by departure delay and print the first few records
flights_tbl %>% filter(dep_delay == 2) %>% head()
```
Let's quickly review some dplyr grammar: The pipe operator %>% indicates feeding the expression on its left-hand side into the function on the right-hand side. This allows us to chain many functions on a dataframe while maintaining readability of the command ( compare the above with head(filter(flights_tbl, dep_delay==2)) ).

A more coplicated example:
```{r}
delay <- flights_tbl %>% 
  group_by(tailnum) %>%
  summarise(count = n(), dist = mean(distance), delay = mean(arr_delay)) %>%
  filter(count > 20, dist < 2000, !is.na(delay)) %>%
  collect
```
Plot the queried data:
```{r}
library(ggplot2)
ggplot(delay, aes(dist, delay)) +
  geom_point(aes(size = count), alpha = 1/2) +
  geom_smooth() +
  scale_size_area(max_size = 2)
```

Now we revie ggplot2 syntax: The first line specifies the x-y axes of the plot. Each additional line preceded by a '+' indicates another layer to be added. The first layer added is a scatter plot where the point size is determined by the counts of instances. The second layer is a smoothed conditional mean with standard errors. The conditional mean is modeled by a generalized additive model (GAM) and the confidence intervals are determined by asymptotics ( I assum;  verify this). We will not take the smoothing function too seriously, it is simply here to roughly illustrate the conditional mean.